---
title: "Day 15 In Class Activity, Part 1"
author: "Vince Melfi"
date: "February 26, 2020"
output:
  html_document:
    number_sections: no
    theme: cerulean
---

```{r global_options, echo = FALSE}
knitr::opts_chunk$set(comment = NA, tidy = TRUE)
```

This ICA focuses on loops. Turn it in to D2L.






## Fibonacci sequence and `for` loops

The Fibonacci sequence is the sequence $\{F_n\colon n \geq 1\}$ of numbers defined by $F_1 = 0, F_2 = 1$ and $F_n = F_{n-2} + F_{n-1}$ for $n\geq 3$. For example the first seven Fibonacci numbers are $F_1 = 0, F_2 = 1$, $F_3 = 0+1 = 1$, $F_4 = 1 + 1 = 2$, $F_5 = 1 + 2 = 3$, $F_6 = 2 + 3 = 5$, and $F_7 = 3 + 5 = 8$. 

**Write a function `Fib` which uses a `for` loop, which takes one argument, `n`, and which returns the first `n` Fibonacci numbers. So for example `Fib1(7)` should return the values `0, 1, 1, 2, 3, 5, 8`. When your function is written, test it by asking for `Fib1(7)` and `Fib1(10)`.** 

```{r}
##Write Fib here and show the output of Fib(7) and Fib(10)
```

Did you take care that your function would return the correct value(s) if called with `Fib(1)` or `Fib(2)`? 

**If so, verify by asking for those values below. If not, add code to your function which will ensure that `Fib(k)` will return the first `k` Fibonacci numbers for positive integers `k` including 0 and 1.**

```{r}
##Either verify that Fib works as expected with arguments n = 1 and n=2, or add code to Fib and then verify that it works. 
```

## Cosines and `while` loops

Here is the cosine function evaluated at the integers 1 through 20:

```{r cos1-20}
cos(1:20)
```

We will next write a function that, given a number $x$ between $-1$ and $1$, calculates the smallest positive integer $k$ for which $\cos(k)$ is within $0.001$ of $x$. From the above evaluation of the cosine function we can see that if $x = -0.653$ then the answer is $k = 4$. 

A little more formally: **Write a function `find_cos_integer` which takes a value $x$ between $-1$ and $1$ (inclusive) as input, and returns the smallest nonnegative integer $k$ satisfying $|\cos(k) - x| < 0.001$. Use a `while` loop. Here are a few examples of such a function in use:**

`> find_cos_integer(1)`    
`[1] 0`  

`> find_cos_integer(0.1)`  
`[1] 7202`

`> find_cos_integer(0.3)`  
`[1] 49`

**You will probably want to include code which stops and returns a message if the user specifies an argument $x$ which is either less than $-1$ or greater than $1$, since otherwise you will likely have an infinite loop on your hands. You will probably find the `abs()` function, as well as the `cos()` function, useful.**

```{r}
##Write find_cos_integer here and test it on arguments 0.1, -0.3, 0.5, 0, 1
```