---
title: "Homework 3"
author: ""
date: "February 21, 2020"
output: 
  html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk
```

The dataset contains information about births in the United States. The full data set is from the Centers for Disease Control. The data for this homework assignment is a "small" sample (chosen at random) of slightly over one million records from the full data set. The data for this homework assignment also only contain a subset of the variables in the full data set. 


## Setting up


```{r echo=FALSE}
library(tidyverse)
library(knitr)
```

```{r echo=FALSE}
birth_data <- read.csv("BirthData.csv", header = TRUE)
birth_data <- as_tibble(birth_data)
```

A glimpse of the data:

```{r }
glimpse(birth_data)
```

The variables in the data set are:

Variable | Description
---------|------------
`year` | the year of the birth
`month` | the month of the birth
`state` | the state where the birth occurred, including "DC" for Washington D.C.
`is_male` | which is `TRUE` if the child is male, `FALSE` otherwise
`weight_pounds` | the child's birth weight in pounds
`mother_age` | the age of the mother    
`child_race` | race of the child. 
`plurality` | the number of children born as a result of the pregnancy, with 1 representing a single birth, 2 representing twins, etc.

## Data structure and wrangling

### Question 1

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


</div>

Here are the states in each region:

##### Northeast Region:

Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island and Vermont, New Jersey, New York, and Pennsylvania

##### Midwest Region:

Illinois, Indiana, Michigan, Ohio and Wisconsin,
Iowa, Kansas, Minnesota, Missouri, Nebraska, North Dakota, and South Dakota

##### South Region:

Delaware, District of Columbia, Florida, Georgia, Maryland, North Carolina, South Carolina, Virginia, and West Virginia,
Alabama, Kentucky, Mississippi, and Tennessee,
Arkansas, Louisiana, Oklahoma, and Texas

##### West Region:

Arizona, Colorado, Idaho, Montana, Nevada, New Mexico, Utah and Wyoming,
Alaska, California, Hawaii, Oregon and Washington

```{r Q1}
birth_data$region <- NA

Northeast <- c("CT","ME","MA","NH","RI","VT","NJ","NY","PA")
Midwest <- c("IL","IN","MI","OH","WI","IA","KS","MN","MO","NE","ND","SD")
South <- c("DE","DC","FL","GA","MD","NC","SC","VA","WV","AL","KY","MS","TN","AR","LA","OK","TX")
West <- c("AZ","CO","ID","MT","NV","NM","UT","WY","AK","CA","HI","OR","WA")

birth_data$region[birth_data$state %in% Northeast] = "NorthEast"
birth_data$region[birth_data$state %in% Midwest] = "Midwest"
birth_data$region[birth_data$state %in% South] = "South"
birth_data$region[birth_data$state %in% West] = "West"
```

### Question 2
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
</div>

##### Red:
Alaska,
Idaho,
Kansas,
Nebraska,
North Dakota,
Oklahoma,
South Dakota,
Utah,
Wyoming,
Texas,
Alabama,
Mississippi,
South Carolina,
Montana,
Georgia,
Missouri,
Louisiana,
Tennessee,
Arkansas,
Kentucky,
Arizona,
West Virginia.

##### Purple:

North Carolina,
Virginia,
Florida,
Ohio,
Colorado,
Nevada,
Indiana,
Iowa,
New Mexico.

##### Blue:

New Hampshire,
Pennsylvania,
California,
Michigan,
Illinois,
Maryland,
Delaware,
New Jersey,
Connecticut,
Vermont,
Maine,
Washington,
Oregon,
Wisconsin,
New York,
Massachusetts,
Rhode Island,
Hawaii,
Minnesota,
District of Columbia.

```{r Q2}
birth_data$state_color <- NA

Red <- c("AK","ID","KS","NE","ND","OK","SD","UT","WY","TX","AL","MS","SC","MT","GA","MO","LA","TN","AR","KY","AZ","WV")
Purple <- c("NC","VA","FL","OH","CO","NV","IN","IA","NM")
Blue <- c("NH","PA","CA","MI","IL","MD","DE","NJ","CT","VT","ME","WA","OR","WI","NY","MA","RI","HI","MN","DC")

birth_data$state_color[birth_data$state %in% Red] = "Red"
birth_data$state_color[birth_data$state %in% Purple] = "Purple"
birth_data$state_color[birth_data$state %in% Blue] = "Blue"
```

Some of the variables have missing values, and these may be related to different data collection choices during different years. For example, possibly plurality wasn't recorded during some years, or state of birth wasn't recorded during some years. In this exercise we investigate using some `dplyr` functions. Hint: The `group_by` and `summarize` functions will help. 

### Question 3

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


</div>

```{r Q3}
data <- read.csv("BirthDataWithRegionColors.csv")


for (i in c(seq(1,length(colnames(data))))){
  cat(colnames(data[i]), ":", sum(is.na(data[i])),"\n" )
}
```

### Question 4

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


</div>

```{r Q4}
data %>% group_by(year) %>%
  summarize(sum(is.na(state)), sum(is.na(child_race)), length(state))
```

### Question 5

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


</div>

```{r Q5}
data %>% group_by(plurality) %>%
  summarize(count = length(weight_pounds), mean_weight= round(mean(weight_pounds,na.rm=TRUE),2), mean_mother_age = round(mean(mother_age),1))
```

### Question 6

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

</div>

```{r Q6}
data %>% group_by(is_male,state_color) %>%
  summarize(count = length(weight_pounds), mean_weight = round(mean(weight_pounds,na.rm=TRUE),2), mean_mother_age = round(mean(mother_age),1))
```

## Data Visualisation:

### Question 1
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

</div>

```{r DV Q1}
attach(data)
mean_info1 <- data %>% group_by(year,region) %>%
  summarize(mean=mean(mother_age,na.rm=TRUE))

ggplot(na.omit(mean_info1),aes(year,mean)) + geom_line(aes(color=region))

```

### Question 2

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


```{r DV Q2}
mean_info2 <- data %>% group_by(year,state_color) %>%
  summarize(mean=mean(mother_age,na.rm=TRUE))
ggplot(na.omit(mean_info2),aes(year,mean)) + geom_line(aes(color = state_color)) + scale_color_manual(values=c("blue", "purple", "red"))
```

### Question 3

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


```{r DV Q3}
mean_weight3 <- data %>% filter(child_race ==  1| child_race == 2) %>% group_by(child_race,year) %>% summarize(mean_weight = mean(weight_pounds, na.rm = TRUE))

ggplot(mean_weight3, aes(year,mean_weight)) + geom_line(aes(color=as.factor(child_race))) + scale_color_discrete(name = "Race of child", labels = c("White","Black"))
```

### Question 4

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

```{r DV Q4}
attach(data)

ggplot(na.omit(data),aes(region,mother_age)) + geom_boxplot()

```
